---
title: "514 Lab 2"
author: "Your Name Here"
date: "Due Date 2/9"
output: html_document
---

```{r message=FALSE}
# If you don't already have the tigerstats library installed,
# you will need to type install.packages("tigerstats") into the Console
library(tigerstats) 
```

### 1. Changing the author field and file name. (5 points)

##### (a) Change the `author:` field on the Rmd document from Your Name Here to your own name.

##### (b) Rename this file to "HW2_YourHameHere.Rmd", where YourNameHere is changed to your own name.

### 2. Measure of location and variability (20 points)

A certain polymer is used for evacuation systems for aircraft. It is important that the polymer be resistant to the aging process. Twenty specimens of the polymer were used in an experiment. Ten were assigned randomly to be exposed to an accelerated batch aging process that involved exposure to high temperatures for 10 days. Measurements of tensile strength of
the specimens were made, and the following data were
recorded on tensile strength in psi:

```
No aging: 227 222 218 216 218 217 225 229 228 221
Aging: 219 214 218 203 215 211 209 204 201 205 
```

```{r}
# You can use the following code to create a data frame
strength = c( 227 ,222, 218, 216, 218, 217, 225, 229, 228,221,219,214,218,203,215,211,209,204,201,205)
aging<-as.factor(c(rep(0,10),rep(1,10)))
polymerData<-data.frame(strength,aging)
```

#### (a) Calculate the sample mean tensile strength of the two samples. (5 points)

```{r}
# Edit me
```

> Type your answer here

#### (b) Calculate the median for both. Discuss your observation with the mean and median of each group. (5 points)

```{r}
# Edit me
```

> Type your answer here

#### (c) Calculate the sample variance as well as standard deviation in tensile strength for both samples. (5 points)

```{r}
# Edit me
```

> Type your answer here

#### (d) Does there appear to be any evidence that aging affects the variability in tensile strength? (5 points)

> Type your answer here

<br> 

### 3. Experiment of Tossing A Die (20 points)

Suppose we have a fair die. Let $n = 10000$ be the number of tossing. 

```{r,fig.width = 6, fig.height = 3}
# To generate a fair die
n <- 10000
x <- c(1:6)
DieToss <- sample(x, size=n, replace=TRUE)
par(mfrow=c(1,2))
barplot(table(DieToss))
hist(DieToss)
```

#### (a) Modify the number of repetition to 10 times, 1000 times, and 100000 times in the handout2-2. Calculate the **mean** of the results instead of sum. (10 points)


```{r}
# Edit me
```


#### (b) Save the results of each repetition and use `par` to display the histogram plot of each repetition. (10 points)


```{r}
# Edit me
```

<br> 

### 4.Normal Distribution with `pnorm` and `qnorm` (20 points)

#### (a) Suppose $Z \sim \mathcal{N}(0,1)$, please use `pnormGC` to find $P(X \le -0.3)$, $P(X \ge 0.5)$, $P( -0.3\le X \le 0.5)$ (10 points)

```{r}
# Edit me
```

#### (b) Please use `qnorm` to find the critical values $z_{0.025}$ and $z_{0.005}$. Which value is larger? Why? (10 points)

```{r}
# Edit me
```

> Type your answer here

### 5. Working With Data (35 points)

* [Obtaining the wine quality  dataset](http://archive.ics.uci.edu/ml/machine-learning-databases/wine-quality/)

#### (a) Import the `winequality-red` data set and name it `winequalRed`. (5 points)

```
# here is a hint for the r version 
# -- change these commands as needed and delete these comments before submitting your work -- 
# if you downloaded the data set as a .csv file then you can read it in as follows: 
# winequalRed <- read.csv("~/Documents/datasets/winequality-red.csv", sep=";")
# To view the data set
#   View(winequalRed)
```


```{r}
# Edit me
```

#### (b) Create a  table of the  `quality` and `alcohol` attributes from the `winequalRed` data set. (5 points)

Do not save the output from the code.
```
# hint: if you have two data columns named X and Y in your data frame, you can use code like this to create  a table: 
table(my.data.set$X, my.data.set$Y)
```

```{r}
# Edit me
```

#### (c) Save the first twenty records of the `winequalRed` data set as their own data frame. (5 points)

```{r}
# Edit me
```

#### (d)  Save the `density` and `pH` records of the `winequalRed` data set as their own data frame. (5 points)

```{r}
# Edit me
```

#### (e) Separate the wine data into a low quality class (quality $\le 5$) and a high quality class (quality > 5), find the mean and standard deviation for the attributes `alcohol` for the two classes. Based on the statistical information, describe if there exists difference for `alcohol` between the low quality and high quality red wines. (10 points)

```{r}
# Edit me
```

> Type your answer here


#### (f) Use `barplot` to investigate `quality` attribute. Show your result. (5 points)

```{r}
# Edit me
```

