---
title: "Homework 5"
author: "Your Name Here"
date: "Due 4/20 11:59 pm"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
   echo = TRUE, 
   fig.align = 'center' , 
   out.width="80%"
)
```

# Setup


### Changing the author field and file name.

* (a) Change the `author:` field on the Rmd document from Your Group Name Here to your group name.

* (b) Rename this file to "hw5_YourNameHere.Rmd", where YourNameHere is changed to your name.

# Cleanup 
Once you are done with this setup, updating the author and date fields, **delete** the *Setup* and the *Cleanup* sections of this document! 

### GRADING

* Part I = 20 points;
* Part II = 80 points;



# Part I: Review of basic concepts in statistical learning (20 points)

You will spend some time thinking of some real-life applications for statistical learning. 

### Question 1. 
Describe three real-life applications in which classification might be useful. Describe the response, as well as the predictors. Is the goal of each application inference or prediction? Explain your answer. 

> Your answer here! 

### Question 2. 
Describe three real-life applications in which regression might be useful. Describe the response, as well as the predictors. Is the goal of each application inference or prediction? Explain your answer.

> Your answer here! 

### Question 3.
Describe three real-life applications in which cluster analysis might be useful.

> Your answer here! 

  
### Question 4. 
What are the advantages and disadvantages of a very flexible (versus a less flexible) approach for regression or classification? Under what circumstances might a more flexible approach be preferred to a less flexible approach? When might a less flexible approach be preferred?

> Your answer here! 

# Part II: Multiple Linear Regression (80 points)

Load the `Boston` data set

```{r}
# import packages
library(MASS)
library(MLmetrics)
#load data
data(Boston)
```


## Exploratory data analysis (10 points)

* Check the number of observations and features using `dim`

```{r}
# Your code here
```

> Your answer here

* Check for missing values

```{r}
# Your code here
```

> Your answer here

* Check for duplicated values

```{r}
# Your code here
```

> Your answer here

* checking correlation between variables 

```{r}
# Your code here
```

> Your answer here


## Split data set into 80:20 train and test data with name `BostonTraining` and `BostonTesting` respectively (10 points)

```{r}
# Your code here
```

> Your answer here


## Subset Selection Linear Regression Model

### Forward Stepwise (25 points)

* Please construct a forward stepwise regression with `BostonTraining`.

```{r}
# Your code here
```

> Your answer here

* Use this model to predict `medv` in `BostonTesting` and calculate `MAE` and `MSE`.

```{r}
# Your code here
```

> Your answer here

### Backward Stepwise (25 points)

* Please construct a backward stepwise regression with `BostonTraining`.

```{r}
# Your code here
```

> Your answer here

* Use this model to predict `medv` in `BostonTesting` and calculate `MAE` and `MSE`.

```{r}
# Your code here
```

> Your answer here

## Model Assessment (10 points)

Compare the forward and backward stepwise linear regression models. You can use plots, assessment measures ($R^2$, RSS, MAE, MSE,etc.) Which one is better? Explain your answer.

> Your answer here

