---
title: "514 Lab 1"
author: "Vikas Sanil"
date: "Due Date 1/26"
output:
  html_document:
    df_print: paged
  pdf_document: default
header-includes:
- \usepackage{float}
- \floatplacement{figure}{H}
fig_cap: yes
keep_tex: yes
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.pos="H")
```
```{r message=FALSE}
library('tidyverse')
```

### 2.  Hello World! (10 points)

```{r }
print("Hello world!")
```

#### (a) Modify the code chunk below to print your name.

```{r }
print("Vikas Sanil")
```

<br>


#### (a) Use `:` to output the sequence of numbers from 3 to 12
```{r 
3:12
```

#### (b) Use `seq()` to output the sequence of numbers from 3 to 30 in increments of 3
```{r }
seq(3,30,3)
```

#### (c) Save the sequence from (a) as a variable `x`, and the sequence from (b) as a variable `y`.  Output their product `x*y`
```{r }
x <- 3:12
y <- seq(3, 30, 3)
x*y
```
\newpage



### 4. Cars data (50 points)

We'll look at data frame and plotting in much more detail in later classes.  For a previous of what's to come, here's a very basic example.

For this example we'll use a very simple dataset.  The `cars` data comes with the default installation of R.  To see the first few columns of the data, just type `head(cars)`.

```{r}
head(cars)
```

#### (a) Calculate the average and standard deviation of `speed` .

```{r }
cat('Average of \'speed\': ', mean(cars$speed))

```

```{r }
cat('Standard Deviation of \'speed\': ', sd(cars$speed))
```

#### (b) Calculate the average and standard deviation of `dist` 

```{r }
cat('Average of \'dist\': ', mean(cars$dist))

```

```{r }
cat('Standard Deviation of \'dist\': ', sd(cars$dist))
```

#### (c)  Produce a histogram of speed using the `qplot` function with **5 bins**.

```{r  out.width="50%", fig.cap="Histogram of speed.", fig.align = 'center', out.extra=''}
qplot(speed,data=cars, bins = 5)
```

\newpage

#### (d) Use the `qplot(x,y)` function to create a scatterplot of dist against speed.

```{r  message=FALSE, out.width="50%", fig.cap="Scatterplot of dist against speed.", fig.align = 'center', fig.pos= 'h'}
qplot(x=speed, y=dist, data=cars, geom = c("point", "smooth"))
```


#### (e) Use the `boxplot` function to create a boxplot of speed.

```{r  out.width="50%", fig.cap="Boxplot of speed.", fig.align = 'center'}
boxplot(cars$speed)
abline(h = min(cars$speed), col = "Blue", )
text(min(cars$speed), paste("Minimum Speed =", min(cars$speed), sep = ' '), adj = c(0.5,-0.1) )
abline(h = max(cars$speed), col = "Yellow")
text(max(cars$speed), paste("Maximum Speed =", max(cars$speed), sep = ' '), adj = c(0.5,-0.1) )
abline(h = median(cars$speed), col = "Green")
text(median(cars$speed), paste("Median Speed =",median(cars$speed), sep = ' '), adj = c(0.5,-0.1) )
abline(h = quantile(cars$speed, c(0.25, 0.75)), col = "Red")
text(quantile(cars$speed, 0.25), paste("25th Percentile of Speed =",quantile(cars$speed, 0.25), sep = ' '), adj = c(0.5,-0.1) )
text(quantile(cars$speed, 0.75), paste("75th Percentile of Speed =",quantile(cars$speed, 0.75), sep = ' '), adj = c(0.5,-0.1) )
```

```{r}
summary(cars$speed)
```

